---
title: "Homework 1"
author: "Narjes Mathlouthi"
date: '2022-04-03'
output: 
  html_document: 
    toc: yes
    code_folding: hide
---

## Machine Learning Main Ideas

Please answer the following questions. Be sure that your solutions are clearly marked and that your document is neatly formatted.

You don’t have to rephrase everything in your own words, but if you quote directly, you should cite whatever materials you use (this can be as simple as “from the lecture/page # of book”).

#### Question 1:

Define supervised and unsupervised learning. What are the difference(s) between them?
- In Supervised learning: (from lecture 1 slides)
Prediction:
• accurately predict future response given predictors
Estimation:
 understand how predictors affect response
Model selection:
• find the “best” model for response given predictors
Inference:
• assess the quality of our predictions and (or) estimation

- In unsupervised learning,the outcome is unknown and where unlabeled data is used to discover hidden patterns in data without the need for human intervention (hence, they are “unsupervised”).

- Supervised learning is the outcome is known vs. unsupervised learning is when the outcome is unknown

#### Question 2:

Explain the difference between a regression model and a classification model, specifically in the context of machine learning.

Machine Learning problems can be characterized into regression or classification models.

- In a regression model, Y is quantitative (or continuous) such as price, blood, pressure (lecture2 slide1)
- In a classification model Y is qualitative (categorical)(e.g., survived/died, spam/not spam) (lecture2 slide1)

#### Question 3:

Name two commonly used metrics for regression ML problems. Name two commonly used metrics for classification ML problems.
Two commonly used metrics for regression ML problems:

- Mean Squared Error (MSE)
- Root Mean Squared Error (RMSE)

Two commonly used metrics for classification ML problems

- Accuracy
- Precision and Recall

#### Question 4:

As discussed, statistical models can be used for different purposes. These purposes can generally be classified into the following three categories. Provide a brief description of each.

Definitions derived from Lecture 2 slide #7

**Descriptive models**:

• What combo of features fits best?
• Aim is to predict Y with minimum reducible error
• Not focused on hypothesis tests

**Inferential models**:

What features are significant?

• Aim is to test theories
• (Possibly) causal claims
• State relationship between outcome & predictor(s)

**Predictive models**:

Choose model to best visually emphasize a trend in data (i.e., using a line on a scatterplot)

#### Question 5:

Predictive models are frequently used in machine learning, and they can usually be described as either mechanistic or empirically-driven. Answer the following questions.

Define mechanistic. Define empirically-driven. How do these model types differ? How are they similar?

- Empirically-driven (or non-parametric) model relies on observations (i.e. no assumptions about f), requires a larger # of observations, is much more flexible and could run into overfitting. (lecture 2 slide 6) 

- A mechanistic model (parametric) uses a theory to predict what will happen in the real world (assume a parametric form
for f). The model won’t match true unknown f, adding more parameters allows for more flexibility.However, too many parameters could result in overfitting.

- A mechanistic model uses theory to predict what will happen in the real world. On the other hand, empirical modeling, studies real-world events (i.e. observations) to develop a theory.

- Both models run into overfitting issues whether it's by adding too many parameters (mechanistic) or by using too many observation (empirical)

In general, is a mechanistic or empirically-driven model easier to understand? Explain your choice.

Mechanistic model assumes a parametric form for f and therefore it's easier to work with. However, it's all depends on the available data we're working with. In some situations, mechanistic models could be easier to use but in models where observations are available it's would be easier to use an empirical model and sometimes both models could be needed to help with understanding the outcome.(lecture2 slides)

Describe how the bias-variance tradeoff is related to the use of mechanistic or empirically-driven models.

The bias tradeoff  of mechanistic or empirically-driven models come with the use of extra parameters that lead to over fitting or using too many observations to over fit the data. This will likely bias the outcome of the models and remove any variance from the models making the models too good to be true.



#### Question 6:

A political candidate’s campaign has collected some detailed voter history data from their constituents. The campaign is interested in two questions:

Given a voter’s profile/data, how likely is it that they will vote in favor of the candidate?

The observations from the voter's profile/data can help us understand their political inclination and help with predicting their voting outcome.

How would a voter’s likelihood of support for the candidate change if they had personal contact with the candidate?

If they had a personal contact with the candidate the voter's likelihood of support will increase significantly 

Classify each question as either predictive or inferential. Explain your reasoning for each.

Te first question uses observations from the voter's profile data to predict their voting outcome.
The second question infers that since the candidate has a relationship with the candidate they will likely vote for him

## Exploratory Data Analysis


This section will ask you to complete several exercises. For this homework assignment, we’ll be working with the mpg data set that is loaded when you load the tidyverse. Make sure you load the tidyverse and any other packages you need.

Exploratory data analysis (or EDA) is not based on a specific set of rules or formulas. It is more of a state of curiosity about data. It’s an iterative process of: generating questions about data visualize and transform your data as necessary to get answers
use what you learned to generate more questions
A couple questions are always useful when you start out. These are “what variation occurs within the variables,” and “what co-variation occurs between the variables.”

You should use the tidyverse and ggplot2 for these exercises.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
```

#### Exercise 1:

We are interested in highway miles per gallon, or the hwy variable. Create a histogram of this variable. Describe what you see/learn.

```{r}
ggplot(mpg, aes(hwy)) + geom_histogram(binwidth = 2) + labs(caption = " ")
```

The histogram seem to be bi modal with two spikes, one spike between 10-20 and another between 20-30.  



#### Exercise 2:

Create a scatterplot. Put hwy on the x-axis and cty on the y-axis. Describe what you notice. Is there a relationship between hwy and cty? What does this mean?

```{r}
ggplot(mpg, aes(x = hwy , y = cty)) +
  geom_point()
```

Highway miles per gallon (hwy) and city miles per gallon (cty) seem to be positively correlated. The miles per gallon increase with more miles which makes sense as drivers use highways with higher mile and cities for lower miles.




#### Exercise 3:

Make a bar plot of manufacturer. Flip it so that the manufacturers are on the y-axis. Order the bars by height. Which manufacturer produced the most cars? Which produced the least?

```{r message=FALSE, warning=FALSE}

ggplot(mpg, aes(x=reorder(manufacturer, manufacturer, function(x)-length(x))))+
  coord_flip()+
  geom_bar(aes(stat = "bin"))
```

Dodge produced the most 
Lincoln produced the least


#### Exercise 4:

Make a box plot of hwy, grouped by cyl. Do you see a pattern? If so, what?

```{r message=FALSE, warning=FALSE}
ggplot(mpg, aes(cyl,hwy, fill=cyl)) +geom_boxplot()
```


On highway, average cyl is 6


#### Exercise 5:

Use the corrplot package to make a lower triangle correlation matrix of the mpg dataset. (Hint: You can find information on the package here.)

Which variables are positively or negatively correlated with which others? Do these relationships make sense to you? Are there any that surprise you?

```{r message=FALSE, warning=FALSE}
library(corrplot)
new_mpg<-select(mpg,displ,cyl,cty, hwy)
M <-cor(new_mpg)
corrplot(M, method="circle", type = "lower")
```

The variables that are positively correlated are hwy & hwy,cty & hwy, cty& cty

The variables that are negatively correlated displ & displ, cyl & displ, cty &displ, hwy & displ, cyl & hwy

**231 Students Only**:


#### Exercise 6:

Recreate the following graphic, as closely as you can. Hint: Use the ggthemes package.

```{r message=FALSE, warning=FALSE}
ggplot(mpg, aes(hwy,class)) +
  geom_boxplot() + 
  geom_jitter(alpha = 0.2)+ 
  theme_minimal() + 
  theme (
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text = element_text(size = 12),
  axis.line.x = element_line(size =0.5),
plot.background = element_rect(size = 0.5))+
  labs(x=" Highway MPG", y= "Vehicle Class")
```

#### Exercise 7:

```{r}
ggplot(mpg, aes(hwy,class, fill= drv)) + coord_flip() +geom_boxplot()
```


#### Exercise 8:

```{r, message=FALSE, warning=FALSE}
ggplot(data=mpg) +
geom_point(mapping = aes(x=displ, y= hwy, color = drv))+
geom_smooth(mapping = aes(x=displ, y= hwy, linetype = drv), se = FALSE)


```





**Citation**

“What Is a Mechanistic Model?” Small Business - Chron.Com, https://smallbusiness.chron.com/mechanistic-model-12706.html. 
